# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
script to read de data set from a zip file

```{r}
w<- read.csv(unz("activity.zip","activity.csv"))
library(ggplot2)
library(lattice)
Sys.setlocale("LC_TIME", "English")

```

Explore and summarize the data

```{r}
str(w)
summary(w)
w$interval<- as.factor(w$interval)
w$date<- as.Date(w$date)
```

agregate the data in days 

```{r}
d<- tapply(w$steps,w$date,na.rm=T, sum)
```

plot the histogram 

```{r}
hist(d, main="Number of steps per day", xlab="steps", col="red",breaks=5)

```


## What is mean total number of steps taken per day?

The mean and the median are: 
```{r}
mean(d, na.rm=T)
median(d,na.rm=T)
````


## What is the average daily activity pattern?

``
the average number of step in the 5 minutes interval is: 
```{r}
five<- tapply(w$steps, w$interval, na.rm=T, mean)
five<- as.data.frame(five)
five<- cbind(as.numeric(rownames(five)),five)
colnames(five)<- c("interval","steps")
````
Graphically the number of stpeps in the 5 minutes intervals.

```{r}
plot(five$interval,five$steps, type="l", 
     main="Mean of number of steps in 5 minutes interval",
     ylab="Steps", xlab="5 minutes interval", col="blue")

```
The five minutes interval with the maximun number of steps is
```{r}
five[five$steps==max(five$steps),1]
````

with the value of 

```{r}
round(max(five$steps),1)
```


## Imputing missing values
The total number of NA's is: 
```{r}
length(w[w$steps=="NA",1])
```
The missing values are replaced with the mean value for 5 minutes interval.A new data frame were created.
first we calculate the mean for the specific interval of time
```{r}
avera5=tapply(w$steps,w$interval,mean, na.rm=T)

```
then, substitute the missing values with the mean of the specific interval of time
```{r}
w2<- w

for(i in 1:17568){
  if (is.na(w2[i,1])){
  w2[i,1]=avera5[w2[i,3]]
  }  
}

```
The new values for mean and median are: 

```{r}
d2<- tapply(w2$steps,w2$date, sum)
hist(d2, main="Number of steps per day", xlab="steps", col="blue",breaks=5)
mean(d2)
median(d2)
````



## Are there differences in activity patterns between weekdays and weekends?

In order to make the two sets of data: 
```{r}

w2$day<-weekdays(w2$date)

wk<- subset(w2, day=="Saturday" | day=="Sunday", select=c(interval, steps))

wkd<- subset(w2,day=="Monday"|day=="Tuesday"|day=="Wednesday" | day=="Thursday"|day=="Friday", select=c(interval, steps))
```

For weekdays:
```{r}
five1<- tapply(wk$steps, wk$interval, mean)
five1<- as.data.frame(five1)
five1<- cbind(as.numeric(rownames(five1)),five1)
colnames(five1)<- c("interval","steps")
````

for weekends:
```{r}
five2<- tapply(wkd$steps, wkd$interval, mean)
five2<- as.data.frame(five2)
five2<- cbind(as.numeric(rownames(five2)),five2)
colnames(five2)<- c("interval","steps")
````
Comparing the two sets of data


```{r}
par(mfrow=c(2,1))
par(mar=c(4,4,2,2))

plot(five2$interval,five2$steps, type="l", 
     main="Weekend",
     xlab="",ylab="Steps", col="blue")

plot(five1$interval,five1$steps, type="l", 
     main="Weekday",
     xlab="Interval",ylab="", col="blue")

```

The graphic shows that the most active period of time during weekends are during the first hours in the morning aproximatly from 8:00 to 10:00 am. 
while during weedays the activity are more variable during the working hours. 